Sparse spectral and $p$-finite element methods on disk-slices and rhombi I: zero boundary conditions

### 2019-01-03 papers/halfdisk/halfdisk.tex
Edits to paper

\include somacros

@@

\sotodoinline{Define norms and inner products below}

\begin{lemma}
$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
\end{align}

@@

Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$\sotodo{Clarify this}. Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e. that $d_2$ is an even integer). Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$. Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
\begin{align}
& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
\end{align}
where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.

@@

\subsection{Heat Equation}
\bstodo{Struggling to get plotting working}
Fix \(a, b = 1,1\), and consider
\begin{align}
\frac{\partial u}{\partial t}(x,y) = \Delta(\Wii \: u)(x,y) \quad \text{in } \Omega
\end{align}

>>>>

\include{somacros}

@@

\begin{lemma}
$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
\end{align}
where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight.

@@

Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e. $f(\cdot, y)$ with $y$ fixed is a degree $d_1$ polynomial, and $f(x, \cdot)$ with $x$ fixed is an even degree $d_2$ polynomial). Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e. that $d_2$ is an even integer). Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$. Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
\begin{align}
& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
\end{align}
where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.

@@

\subsection{Heat Equation}
Fix \(a, b = 1,1\), and consider
\begin{align}
\frac{\partial u}{\partial t}(x,y) = \Delta(\Wii \: u)(x,y) \quad \text{in } \Omega
\end{align}
